Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
work on spec model; fixing figure 7

@@

Similarly to the single-entry vector, zero-entry vectors 
are also indexing schemes which generalising all other schemes
which at least have the $r^{th}$ entry is $0$, representing indices at
the ``origin'' in dimension $r$. 
\end{definition}

\begin{figure}
\input{region-model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Model of regions, $\interp{\textit{region}}$}
\label{fig:region-model}
\end{figure}

@@

The intuition behind $\otimes$ is that it takes
all possible pairs of indexing schemes, treating them as the
lower and upper bounds of an $n$-dimensional rectangle from which the 
remaining vertices of the rectangle are generated (like a bounding
box). \eg{}:
%%
\begin{align*}
\{\vtwoh{1}{2}, \vtwoh{5}{6}\} \otimes \{\vtwoh{3}{4}\} = 
\{&\vtwoh{1}{2}, \vtwoh{3}{2}, \vtwoh{1}{4}, \vtwoh{3}{4}, \\
  & \vtwoh{3}{6}, \vtwoh{5}{4}, \vtwoh{5}{6}\}
\end{align*}
%%
Internally $\otimes$ takes a Cartesian product of the indexing
schemes (all pairs) and then combines each pair using the 
\emph{pairwise permutation} written $\bowtie$ (defined later)
which generates all possible interleavings of two vectors
(\ie{}, all possible non-deterministic choices between $u_i$ and
$v_i$ in each dimension). Some care is then taken over
how this interacts with $\infty$: pariwse permtuation
corresponds to non-determinstic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either 
is $\infty$ then the resultying pairwise permutation is 
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} = 
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
% 
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.
%With respect to the notion of unification, the 
%
%
% combines non-$\infty$ constraints
%of the unifiers in different dimensions into less accepting unifiers.
%This means an access pattern generalises with this scheme would require
%both models to be satisfied. Use of $dims$ gives the dimensions that
%constitutent models talk about and is used in the product to eliminate
%general unifier appearing in these dimensions. Allowing general unifier
%in these dimensions would prevent both models to be satisfied
%simultaneously.
%  u in M * N   u < w
% mplies that 
%  u_1 in M  u1 < w
%  u_2 in M  u2 < w
%
The notion of which dimensioons $i$ have that \emph{any scheme in the model
has $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
%
\begin{definition}[Constrained dimensions]
Given an $n$-dimensional model $M$, the \emph{constrained dimensions}
of that model is a subset of $\{1, \ldots, n\}$ of 
the dimensions for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}

@@

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\ 
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
% \\
%\setlength{\arraycolsep}{0.5em}
%\left[\begin{array}{ccc}
%0 & 1 & 2 \\
%0 & 1 & 5 \\
%0 & 4 & 2 \\
%0 & 4 & 5 \\
%3 & 1 & 2 \\
%3 & 1 & 5 \\
%3 & 4 & 2 \\
%3 & 4 & 5
%\end{array}
%\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\end{definition}

@@

\begin{figure}
\input{spatial-model.tex}
\caption{Model of specifications, $\interp{\textit{spec}}$}
\label{fig:spatial-model}
\end{figure}

@@

\todo{Explain the injections and the domain here}

@@

\begin{figure}[t]
\vspace{-0.3em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0.2em]
{\scalebox{0.75}{
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} 
}} \\[0.36em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
{\scalebox{0.75}{
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture}
}}
 \\ (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
{\scalebox{0.75}{
\begin{tikzpicture}
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture}
}}  \\ (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in 
specification inference}
\label{fig:inference-steps-informal}
\end{figure}

>>>>

\usetikzlibrary{calc}

%% http://tex.stackexchange.com/questions/55068/is-there-a-tikz-equivalent-to-the-pstricks-ncbar-command
\tikzset{
    ncbar angle/.initial=90,
    ncbar/.style={
        to path=(\tikztostart)
        -- ($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)
        -- ($(\tikztotarget)!($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztostart)$)
        -- (\tikztotarget)
    },
    ncbar/.default=0.5cm,
}

@@

Similarly to the single-entry vector, zero-entry vectors 
are generalise all other schemes
which at least have the $r^{th}$ entry is $0$, representing indices at
the ``origin'' in dimension $r$. 
\end{definition}

\begin{figure}
\input{region-model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Model of regions, 
$\interp{-}_n : \textit{region} \rightarrow \mathcal{P}(\domainVal{}^n)$}
\label{fig:region-model}
\end{figure}

@@

The intuition behind $\otimes$ is that it takes
all possible pairs of indexing schemes, treating them as the
lower and upper bounds of an $n$-dimensional rectangle from which the 
remaining vertices of the rectangle are generated (like a bounding
box). \eg{}:
%%
\begin{align*}
\{\vtwoh{1}{2}, \vtwoh{5}{6}\} \otimes \{\vtwoh{3}{4}\} = 
\{&\vtwoh{1}{2}, \vtwoh{3}{2}, \vtwoh{1}{4}, \vtwoh{3}{4}, \\
  & \vtwoh{3}{6}, \vtwoh{5}{4}, \vtwoh{5}{6}\}
\end{align*}
%%
The tensor takes a Cartesian product of the two models,
giving all pairs of schemes which are combined using the 
\emph{pairwise permutation} written $\bowtie$ (defined later).
Pairwise permutation generates all possible interleavings of two vectors
(\ie{}, all possible non-deterministic choices between $u_i$ and
$v_i$ in each dimension). Some care is then taken over
how this interacts with $\infty$: pairwise permtuation
corresponds to non-determinstic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either 
is $\infty$ then the resultying pairwise permutation is 
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} = 
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
% 
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.
%With respect to the notion of unification, the 
%
%
% combines non-$\infty$ constraints
%of the unifiers in different dimensions into less accepting unifiers.
%This means an access pattern generalises with this scheme would require
%both models to be satisfied. Use of $dims$ gives the dimensions that
%constitutent models talk about and is used in the product to eliminate
%general unifier appearing in these dimensions. Allowing general unifier
%in these dimensions would prevent both models to be satisfied
%simultaneously.
%  u in M * N   u < w
% mplies that 
%  u_1 in M  u1 < w
%  u_2 in M  u2 < w
%
The notion of which dimensioons $i$ have that \emph{any scheme in the model
has $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
%
\begin{definition}[Constrained dimensions]
Given an $n$-dimensional model $M$, the \emph{constrained dimensions}
of that model is a subset of $\{1, \ldots, n\}$ of 
the dimensions for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}

@@

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\ 
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
% \\
%\setlength{\arraycolsep}{0.5em}
%\left[\begin{array}{ccc}
%0 & 1 & 2 \\
%0 & 1 & 5 \\
%0 & 4 & 2 \\
%0 & 4 & 5 \\
%3 & 1 & 2 \\
%3 & 1 & 5 \\
%3 & 4 & 2 \\
%3 & 4 & 5
%\end{array}
%\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. The pairwise permutation
matrix  s thus defined in terms of the logical
matrix $b$ where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$ as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
\end{definition}

@@

@@

Thus far we have explained the model of regions as sets of index
schemes, that is
$\interp{\textit{region}}_n \in \mathcal{P}(\domainVal{}^n)$.
Figure~\ref{fig:spatial-model} defines the top-level model of the
\textit{spec} syntax, which includes the multiplicity and
approximation modifiers.

\begin{figure}
\input{spatial-model.tex}
\caption{\textit{spec} model, 
$\interp{-}_n : \textit{spec} \rightarrow \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$}
\label{fig:spatial-model}
\end{figure}

@@

\begin{figure}[t]
\vspace{-0.3em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0.2em]
\begin{tikzpicture}[scale=0.75]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0.36em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.75]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\ (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.75]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\ (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in 
specification inference}
\label{fig:inference-steps-informal}
\end{figure}
